Assessment of a Shen-Fourier turbulent channel flow solver for large-scale simulations

### 2016-09-20 paper.tex
Added MHD
Added a new section containing MHD discretization, and extended the appendix by including matrices that are used in the MHD section.
Some speed results

If the expansion coefficients $\hat{u}$  are known for the entire wavenumber 
mesh (\ref{eq:wavenumbermesh}), then the expression (\ref{eq:u_solx}) may be 
evaluated on the mesh (\ref{eq:Xn}) using fast Fourier and fast Shen inverse 
transforms as
\begin{align}
u(x_i, y_j, z_k, t) &= \underbrace{\frac{1}{N_z}\sum_n 
\underbrace{\frac{1}{N_y} \sum_m \underbrace{\sum_l 
\hat{u}(l,m,n)\N{\phi}_l(x_i)}_{\N{\mathcal{S}}_x^{-1}} e^{\imath \underline{m} 
y_j}}_{\mathcal{F}_y^{-1}} e^{\imath \underline{n} z_k}}_{\mathcal{F}_z^{-1}} , 
\quad \forall \, \bm{x} \in X_N, \notag \\
  u &= \N{\mathcal{T}}^{-1}(\hat{u}) =  \mathcal{F}^{-1}_z (\mathcal{F}^{-1}_y 
  (\N{\mathcal{S}}^{-1}_x (\hat{u}))), \quad \forall \, \bm{x} \in X_N.  
  \label{eq:ifft} 
\end{align}
Here $\N{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse 
transform, $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represent inverse 
Fourier transforms along directions $y$ and $z$ respectively, and the inverse 
Shen transform $\N{\mathcal{S}}_{x}^{-1}$ is performed along the wall normal 
$x$ direction. Note that the transforms are applied sequentially to the entire 
computational mesh. The inverse Shen transform may be computed using fast 
Chebyshev transforms for all the three bases in (\ref{eq:Tk}, \ref{eq:phiD}, 
\ref{eq:phiN}). The transforms are slightly different for the three bases and 
${\mathcal{S}}, \D{\mathcal{S}}$ and $\N{\mathcal{S}}$ are used to distinguish 
between them with obvious notation.

A three-dimensional scalar product in the weighted $[L^2_w(\Omega)]^3$ space is defined as
\begin{align}
 \left<u, v\right>_w &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}w\,dxdydz, \notag \\
   &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}\,dx_wdydz,
\end{align}
where $v^*$ is the comlex conjugate of $v$ and the weights $w$ are unity for periodic directions and  $w=1/\sqrt{1-x^2}$ for the inhomogeneous direction, and we have used $dx_w = dx/\sqrt{1-x^2}$. In this work we will make use of the discrete weighted $[l^2_w(\Omega)]^3$ space, where quadrature is employed for integration. As such the scalar product corresponds to sums that may be computed using fast transforms
\begin{align}
 \left< u, \N{\psi}_{\bm{k}} \right>_w &= h \underbrace{\sum_i 
 \underbrace{\sum_j \underbrace{\sum_k u(x_i, y_j, z_k, t)  e^{-\imath 
 \underline{n} z_k}}_{\mathcal{F}_n}  e^{-\imath \underline{m} y_j} 
 }_{\mathcal{F}_m} \N{\phi}_{l}(x_i) w_i}_{\N{\mathcal{S}}_l},   \notag \\
  &=  h \N{\mathcal{S}}(u) = h \N{\mathcal{S}}_{l} (\mathcal{F}_{{m}} 
  (\mathcal{F}_{{n}}(u))), \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:sst1}
\end{align}
Here $h\N{\mathcal{S}}$ denotes the complete three dimensional scalar product 
and $h = L_yL_zN_y^{-1}N_z^{-1}$ is a constant. $\mathcal{F}_{{n}}$ and 
$\mathcal{F}_{{m}}$ represent discrete Fourier transforms in $z$- and 
$y$-directions, respectively, and $\N{\mathcal{S}}_{l}(\cdot) = (\cdot, 
\N{\phi}_l)_w$ is the Shen scalar product in the $x$-direction. The weights, 
$w_i$, required for the Shen transforms are 
defined as
\begin{equation}
 w_i = \begin{cases}
       \frac{\pi}{c_i N_x} &\forall \, i=0,1,\ldots, N_x \quad  \text{for 
       Gauss-Lobatto points},\\
       \frac{\pi}{N_x+1} &\forall \, i=0,1,\ldots, N_x  \quad \text{for 
       Gauss-Chebyshev points},      
 \end{cases}
\end{equation}
where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$.

The scalar product  $h\N{\mathcal{S}}$ in (\ref{eq:sst1}) does not represent a 
complete transform. To find a transformation between physical $u$ and spectral 
$\hat{u}$ we make use of Eq. (\ref{eq:u_solx}) directly and use the discrete 
orthogonality of the Fourier basis functions
\begin{align}
\left<u, \N{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< 
\sum_{\bm{m}=(q,r,s) \in \N{K}_N} \hat{u}_{\bm{m}} \N{\psi}_{\bm{m}}, 
\N{\psi}_{\bm{k}} \right>_w, \notag \\
           &= h \sum_{q=0}^{N_x-4} \sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) 
           \N{\phi}_{l}(x_i) w_i \, \hat{u}(q, {m}, {n}, t), \notag \\
           &= h \sum_{q=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t), 
           \label{eq:sst2}
\end{align}
where $\N{B}_{lq} = (\N{\phi}_q, \N{\phi}_l)_w = 
\sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) \N{\phi}_{l}(x_i) w_i$ is a 
pentadiagonal mass matrix. The complete transformation is now obtained by 
setting Eq. (\ref{eq:sst2}) equal to (\ref{eq:sst1}) and solving for $\hat{u}$. 
Moving to matrix notation we get
\begin{align}
h\sum_{q=0}^{N_x-4} \N{B}_{lq}\, \hat{u}(q, {m}, {n}, t) &= 
h\N{\mathcal{S}}(u)(l, m, n, t), \notag \\
 \hat{u}_{\bm{k}}(t) &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u), 
 \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:fstB}
\end{align}
where $\N{\mathcal{T}}(u)$ denotes the complete transformation, such that $u = 
\N{\mathcal{T}}^{-1}(\N{\mathcal{T}}(u))$. Note that since $\N{B}$ assembles to 
a pentadiagonal matrix the solution ($\N{B}^{-1}$) can be obtained very fast 
and 
the complete 
transformation thus requires a fast Chebyshev transform ($O(N_x \log N_x)$) and 
a fast linear algebraic solve ($O(N_x)$). Similar transforms $\mathcal{T}$ 
and $\D{\mathcal{T}}$ are defined for the two other bases (\ref{eq:Tk}) and 
(\ref{eq:phiD}), using mass matrices $B_{lq}=(T_q, T_l)_w$ (diagonal) and 
$\D{B}_{lq}=(\D{\phi}_l, \D{\phi}_q)_w$ (tridiagonal) and scalar products 
$\mathcal{S}_l(\cdot) = (\cdot, T_l)_w$ and $\D{\mathcal{S}}_l(\cdot) = (\cdot, 
\D{\phi}_l)_w$.

\section{Discretization of Navier Stokes equations}
\label{sec:discretizationNS}
The Navier Stokes equations (\ref{eq:NS}) are solved using a scheme 
proposed by Kim, Moin and Moser (1987). This scheme is developed by taking the 
Laplacian of the wall normal momentum equation and the curl of the momentum 
equation. Following elimination of the pressure, the equations to solve are 
\begin{align}
\frac{\partial}{\partial t} \nabla^2 u &= h_u + \nu \nabla^4 u, 
\label{eq:biharmonic} \\
\frac{\partial g}{\partial t} &= h_g + \nu \nabla^2 g, \label{eq:g} \\
f + \frac{\partial u}{\partial x} &= 0, \label{eq:f}
\end{align}
where
\begin{align}
f &= \frac{\partial v}{\partial y} + \frac{\partial w}{\partial z}, \\
g &= \frac{\partial w}{\partial y} - \frac{\partial v}{\partial z}, \\
h_u = -\frac{\partial}{\partial x} &\left( \frac{\partial 
\mathcal{H}_y}{\partial y} + \frac{\partial \mathcal{H}_z}{\partial z} \right) 
+ \left(\frac{\partial^2}{\partial y^2} + \frac{\partial^2}{\partial z^2} 
\right) \mathcal{H}_x ,
\\
h_g &= \frac{\partial \mathcal{H}_z}{\partial y} - \frac{\partial 
\mathcal{H}_y}{\partial z}.
\end{align}
The two remaining velocity components are computed from the definitions of $f$ 
and $g$. The biharmonic equation (\ref{eq:biharmonic}) is solved with boundary 
conditions are $u(\pm 1) = u'(\pm 1) = 0$, where the Neumann conditions follow 
from the continuity equation. The boundary conditions for Eq. (\ref{eq:g}) are 
$g(\pm 1) = 0$. The variational formulation becomes: Find ${u} \in 
\N{V}_N$, ${g} \in \D{V}_N$ and ${f} \in \D{V}_N$ such that
\begin{align}
	\frac{\partial }{\partial t} \left< \nabla^2 u, \N{\psi}\right>_w &= 
	\left<h_u, \N{\psi} \right>_w + \nu \left<\nabla^4u, \N{\psi}\right>_w 
	&\forall \N{\psi} \in \N{V}_N, \label{eq:u1} \\
	\frac{\partial}{\partial t}\left<g, \D{\psi}\right>_w &= \left<h_g, 
	\D{\psi}\right>_w + \nu 
	\left<\nabla^2 g, \D{\psi}\right>_w &\forall \D{\psi} \in \D{V}_N ,
	\label{eq:g1} \\
	\left<f, \D{\psi}\right>_w &= \left<\frac{\partial u}{\partial x}, 
	\D{\psi}\right>_w &\forall \D{\psi} \in \D{V}_N. \label{eq:f1}
\end{align}
The two remaining velocity components are computed by projection to the 
Dirichlet space $\D{V}_N$: Find 
${v}, {w}$ in $\D{V}_N \times \D{V}_N$ such that
\begin{align}
\left<f, \D{\psi}\right>_w &= \left<\frac{\partial v}{\partial y} + 
\frac{\partial w}{\partial z}, \D{\psi}\right>_w \, &\forall \D{\psi} \in 
\D{V}_N, \label{eq:f2} \\
\left<g, \D{\psi}\right>_w &= \left<\frac{\partial w}{\partial y}  - 
\frac{\partial v}{\partial z}, \D{\psi}\right>_w \, &\forall \D{\psi} \in 
\D{V}_N, \label{eq:g2}
\end{align}
which simplifies considerably because all the derivatives are in periodic 
directions. Written in spectral space Eqs. (\ref{eq:f2}) and (\ref{eq:g2}) 
become simply
\begin{align}
\hat{f}_{\bm{k}} &= \imath \underline{m}\, \hat{v}_{\bm{k}} + \imath 
\underline{n}\, \hat{w}_{\bm{k}} &\forall \bm{k} \in \D{K}_N^0, \label{eq:f3} \\
\hat{g}_{\bm{k}} &= \imath \underline{m}\, \hat{w}_{\bm{k}} - \imath 
\underline{n}\, \hat{v}_{\bm{k}} & \forall \bm{k} \in \D{K}_N^0, \label{eq:g3}
\end{align}
where $\D{K}_N^0$ is used to denote that these equations can be solved for all 
wavenumbers except $m=n=0$. For $m=n=0$ we solve instead the 
corresponding momentum equations in $y$ and $z$ directions:
\begin{align}
\frac{\partial }{\partial t} \D{B}_{lq}\hat{v}(q, 0, 0, t) &= 
\D{B}_{lq}\hat{\mathcal{H}}_y(q, 0, 0, t) + \nu \D{A}_{lq} \hat{v}(q, 0, 0, t) 
& \forall\, l \in [0, 1, 
\ldots, N_x-2], \label{eq:v0}\\
\frac{\partial }{\partial t} \D{B}_{lq}\hat{w}(q, 0, 0, t) &= 
\D{B}_{lq}\hat{\mathcal{H}}_z(q, 0, 0, t) + \nu \D{A}_{lq} \hat{w}(q, 0, 0, t) 
& \forall\, l \in [0, 1, \ldots, N_x-2].\label{eq:w0}
\end{align}

Equations (\ref{eq:u1}-\ref{eq:f1}) are also solved in spectral space. 
Inserting for (\ref{eq:u_solx}) and a similar expansion for $g$ in $\D{V}_N$, 
the required inner products become
\begin{align}
\left<\nabla^4u, \N{\psi}_{\bm{k}}\right>_w &= h\left[ \left( 
\N{\phi}_q^{''''}, 
\N{\phi}_l\right)_w -2(\underline{m}^2+\underline{n}^2) \left( \N{\phi}_q^{''}, 
\N{\phi}_l\right)_w + (\underline{m}^2+\underline{n}^2)^2\left( \N{\phi}_q, 
\N{\phi}_l\right)_w  \right] \hat{u}_q, \\
\left< \nabla^2 u, \N{\psi}_{\bm{k}}\right>_w &= h\left[\left( \N{\phi}_q^{''}, 
\N{\phi}_l\right)_w - (\underline{m}^2+\underline{n}^2)\left( \N{\phi}_q, 
\N{\phi}_l \right)_w \right] \hat{u}_q, \\
\left< \nabla^2 g, \D{\psi}_{\bm{k}}\right>_w &= h\left[\left( \D{\phi}_q^{''}, 
\D{\phi}_l\right)_w - (\underline{m}^2+\underline{n}^2)\left( \D{\phi}_q, 
\D{\phi}_l \right)_w \right] \hat{g}_q, \\
\left<\frac{\partial u}{\partial x}, \D{\psi}_{\bm{k}}\right>_w &=
h\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w \hat{u}_q, \\
\left<h_u, \N{\psi} \right>_w &= h\left[-(\underline{m}^2+\underline{n}^2) 
\N{\mathcal{S}}(\mathcal{H}_x) - \imath \underline{m}\left(\D{\phi}_q^{'}, 
\N{\phi}_l \right)_w \hat{\mathcal{H}}_{y, q} - \imath 
\underline{n}\left(\D{\phi}_q^{'}, \N{\phi}_l \right)_w 
\hat{\mathcal{H}}_{z, q}\right], 
\\
\left< h_g, \D{\psi} \right>_w &= h\left[ \imath \underline{m}\, \D{B} 
\hat{\mathcal{H}}_z - \imath \underline{n}\, \D{B} \hat{\mathcal{H}}_{y} 
\right],
\end{align}
where for brevity in notation (as before) it is understood that the matrices 
act along the first dimension of the transformed variables, i.e.,  $ 
\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w \hat{u}_q$ is short for the matrix 
vector product $ \sum_{q=0}^{N_x-4}\left( \N{\phi}_q^{'}, \D{\phi}_l\right)_w 
\hat{u}(q, {m}, {n}, t)$. The inner products are used to set up linear systems 
of equations for the inhomogeneous wall normal direction. All inner products 
$\left(\cdot, \cdot\right)_w$ can be represented by sparse matrices.
%\begin{align}
%\frac{\partial }{\partial t} \left( \N{\psi}_j^{''}, \N{\psi}_k\right)_w 
%\hat{u}_j &=  \left(h_u, \N{\psi}_k \right)_w + \nu \left(\N{\psi}^{''''}_j, 
%\N{\psi}_k\right)_w \hat{u}_j, \, &\forall \N{\psi}_k \in \N{V}_N, \\
%\frac{\partial}{\partial t}\left(\D{\psi}_j, \D{\psi}_k\right)_w \hat{g}_j &= 
%\left(h_g, \D{\psi}_k\right)_w + \nu \left(\D{\psi}_j^{''}, 
%\D{\psi}_k\right)_w 
%\hat{g}_j, \, &\forall \D{\psi}_k \in \D{V}_N \\
%\left(\D{\psi}_j, \D{\psi}_k\right)_w \hat{f}_j &= \left(\N{\psi}_j^{'}, 
%\D{\psi}_k\right)_w \hat{u}_j, \, &\forall \D{\psi}_k \in \D{V}_N.
%\end{align}
The mass matrices, stiffness matrices $\N{A}_{lq} = \left( \N{\phi}_q^{''}, 
\N{\phi}_l\right)_w, 
\D{A}_{lq} = \left( \D{\phi}_q^{''}, \D{\phi}_l\right)_w 
$ and $\N{Q}_{lq} = \left(\N{\phi}^{''''}_q, \N{\phi}_l\right)_w$ are given by 
Shen, whereas $\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w$ and 
$\left(\D{\phi}_q^{'}, \N{\phi}_l\right)_w$ are given in the 
appendix.

>>>>

If the expansion coefficients $\hat{u}$  are known for the entire wavenumber 
mesh (\ref{eq:wavenumbermesh}), then the expression (\ref{eq:u_solx}) may be 
evaluated on the mesh (\ref{eq:Xn}) using fast Fourier and fast Shen inverse 
transforms as
\begin{align}
u(x_i, y_j, z_k, t) &= \underbrace{\frac{1}{N_z}\sum_n 
\underbrace{\frac{1}{N_y} \sum_m \underbrace{\sum_l 
\hat{u}(l,m,n,t)\N{\phi}_l(x_i)}_{\N{\mathcal{S}}_x^{-1}} e^{\imath 
\underline{m} 
y_j}}_{\mathcal{F}_y^{-1}} e^{\imath \underline{n} z_k}}_{\mathcal{F}_z^{-1}} , 
\quad \forall \, \bm{x} \in X_N, \notag \\
  u &= \N{\mathcal{T}}^{-1}(\hat{u}) =  \mathcal{F}^{-1}_z (\mathcal{F}^{-1}_y 
  (\N{\mathcal{S}}^{-1}_x (\hat{u}))), \quad \forall \, \bm{x} \in X_N.  
  \label{eq:ifft} 
\end{align}
Here $\N{\mathcal{T}}^{-1}$ is used as short notation for the complete inverse 
transform, $\mathcal{F}_{y}^{-1}$ and $\mathcal{F}_{z}^{-1}$ represent inverse 
Fourier transforms along directions $y$ and $z$ respectively, and the inverse 
Shen transform $\N{\mathcal{S}}_{x}^{-1}$ is performed along the wall normal 
$x$ direction. Note that the transforms are applied sequentially to the entire 
computational mesh. The inverse Shen transform may be computed using fast 
Chebyshev transforms for all the three bases in (\ref{eq:Tk}, \ref{eq:phiD}, 
\ref{eq:phiN}). The inverse transforms are slightly different for the three 
bases and 
${\mathcal{S}}^{-1}, \D{\mathcal{S}}^{-1}$ and $\N{\mathcal{S}}^{-1}$ are used 
to distinguish between them with obvious notation.

A three-dimensional scalar product in the weighted $[L^2_w(\Omega)]^3$ space is defined as
\begin{align}
 \left<u, v\right>_w &= \int_{\Omega} {u(\bm{x}) v^*(\bm{x})}w\,dxdydz, 
\end{align}
where $v^*$ is the comlex conjugate of $v$ and the weights $w$ are unity for 
periodic directions and  $w=1/\sqrt{1-x^2}$ for the inhomogeneous direction. In 
this work we will make use of the discrete weighted $[l^2_w(\Omega)]^3$ space, 
where quadrature is employed for integration. As such the scalar product 
corresponds to sums that may be computed using fast transforms. Using 
$\N{\psi}_{\bm{k}}$ for $v$ we obtain
\begin{align}
 \left< u, \N{\psi}_{\bm{k}} \right>_w &= h \underbrace{\sum_i 
 \underbrace{\sum_j \underbrace{\sum_k u(x_i, y_j, z_k, t)  e^{-\imath 
 \underline{n} z_k}}_{\mathcal{F}_n}  e^{-\imath \underline{m} y_j} 
 }_{\mathcal{F}_m} \N{\phi}_{l}(x_i) w_i}_{\N{\mathcal{S}}_l},   \notag \\
  &=  h \N{\mathcal{S}}(u) = h \N{\mathcal{S}}_{l} (\mathcal{F}_{{m}} 
  (\mathcal{F}_{{n}}(u))), \quad \forall \, \bm{k} \in \N{K}_N. \label{eq:sst1}
\end{align}
Here $h\N{\mathcal{S}}$ denotes the complete three dimensional scalar product 
and $h = L_yL_zN_y^{-1}N_z^{-1}$ is a constant. $\mathcal{F}_{{n}}$ and 
$\mathcal{F}_{{m}}$ represent discrete Fourier transforms in $z$- and 
$y$-directions, respectively, and $\N{\mathcal{S}}_{l}(\cdot) = (\cdot, 
\N{\phi}_l)_w$ is used to represent the Shen scalar product in the 
$x$-direction. The weights, $w_i$, required for the Shen transforms are 
defined as
\begin{equation}
 w_i = \begin{cases}
       \frac{\pi}{c_i N_x} &\forall \, i=0,1,\ldots, N_x \quad  \text{for 
       Gauss-Lobatto points},\\
       \frac{\pi}{N_x+1} &\forall \, i=0,1,\ldots, N_x  \quad \text{for 
       Gauss-Chebyshev points},      
 \end{cases}
\end{equation}
where $c_0 = c_{N_x} = 2$ and $c_i = 1$ for $0 < i < N_x$.\footnote{Note that 
$c_{N_x}=1$ for the continuous scalar product $L^2_w(\Omega)$ (see, e.g., Eq. 
(2.7) of Shen2), but for the discrete $l^2_w(\Omega)$ it follows that 
$c_{N_x}=2$ due to inexact quadrature for the highest wavenumber.}

The scalar product  $h\N{\mathcal{S}}$ in (\ref{eq:sst1}) does not represent a 
complete transform. To find a transformation between physical $u$ and spectral 
$\hat{u}$ we make use of Eq. (\ref{eq:u_solx}) directly and use the discrete 
orthogonality of the Fourier basis functions
\begin{align}
%\left<u, \N{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< 
%\sum_{\bm{m}=(q,r,s) \in \N{K}_N} \hat{u}_{\bm{m}} \N{\psi}_{\bm{m}}, 
%\N{\psi}_{\bm{k}} \right>_w, \notag \\
\left<u, \N{\psi}_{\bm{k}}\right>_w &= \frac{1}{N_yN_z}\left< 
\sum_{(q,r,s) \in \N{K}_N} \hat{u}(q,r,s,t) \N{\psi}(q,r,s), 
\N{\psi}(l,m,n,t) \right>_w, \notag \\
           &= h \sum_{q=0}^{N_x-4} \sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) 
           \N{\phi}_{l}(x_i) w_i \, \hat{u}(q, {m}, {n}, t), \notag \\
           &= h \sum_{q=0}^{N_x-4} \N{B}_{lq} \hat{u}(q, {m}, {n}, t), 
           \label{eq:sst2}
\end{align}
where $\N{B}_{lq} = (\N{\phi}_q, \N{\phi}_l)_w = 
\sum_{i=0}^{N_x} \N{\phi}_{q}(x_i) \N{\phi}_{l}(x_i) w_i$ is a 
pentadiagonal mass matrix. The complete transformation is now obtained by 
setting Eq. (\ref{eq:sst2}) equal to (\ref{eq:sst1}) and solving for $\hat{u}$. 
Moving to matrix notation we get
\begin{align}
h\sum_{q=0}^{N_x-4} \N{B}_{lq}\, \hat{u}(q, {m}, {n}, t) &= 
h\N{\mathcal{S}}(u)(l, m, n, t), \notag \\
 \hat{u}_{\bm{k}}(t) &= \N{\mathcal{T}}(u) =  \N{B}^{-1}\N{\mathcal{S}}(u), 
 \quad \forall \, \bm{k} \in \N{K}_N, \label{eq:fstB}
\end{align}
where $\N{\mathcal{T}}(u)$ denotes the complete transformation, such that $u = 
\N{\mathcal{T}}^{-1}(\N{\mathcal{T}}(u))$. Note that since $\N{B}$ assembles to 
a pentadiagonal matrix the solution ($\N{B}^{-1}$) can be obtained very fast 
and 
the complete 
transformation thus requires a fast Chebyshev transform ($O(N_x \log N_x)$) and 
a fast linear algebraic solve ($O(N_x)$). Similar transforms $\mathcal{T}$ 
and $\D{\mathcal{T}}$ are defined for the two other bases (\ref{eq:Tk}) and 
(\ref{eq:phiD}), using mass matrices $B_{lq}=(T_q, T_l)_w$ (diagonal) and 
$\D{B}_{lq}=(\D{\phi}_l, \D{\phi}_q)_w$ (tridiagonal) and scalar products 
$\mathcal{S}_l(\cdot) = (\cdot, T_l)_w$ and $\D{\mathcal{S}}_l(\cdot) = (\cdot, 
\D{\phi}_l)_w$.

\section{Discretization of Navier Stokes equations}
\label{sec:discretizationNS}
The Navier Stokes equations (\ref{eq:NS}) are solved using a scheme 
proposed by Kim, Moin and Moser (1987). This scheme is developed by taking the 
Laplacian of the wall normal momentum equation and the curl of the momentum 
equation. Following elimination of the pressure, the equations to solve are 
\begin{align}
\frac{\partial}{\partial t} \nabla^2 u &= h_u + \nu \nabla^4 u, 
\label{eq:biharmonic} \\
\frac{\partial g}{\partial t} &= h_g + \nu \nabla^2 g, \label{eq:g} \\
f + \frac{\partial u}{\partial x} &= 0, \label{eq:f}
\end{align}
where
\begin{align}
f &= \frac{\partial v}{\partial y} + \frac{\partial w}{\partial z}, \\
g &= \frac{\partial w}{\partial y} - \frac{\partial v}{\partial z}, \\
h_u = -\frac{\partial}{\partial x} &\left( \frac{\partial 
\mathcal{H}_y}{\partial y} + \frac{\partial \mathcal{H}_z}{\partial z} \right) 
+ \left(\frac{\partial^2}{\partial y^2} + \frac{\partial^2}{\partial z^2} 
\right) \mathcal{H}_x ,
\\
h_g &= \frac{\partial \mathcal{H}_z}{\partial y} - \frac{\partial 
\mathcal{H}_y}{\partial z}.
\end{align}
The two remaining velocity components are computed from the definitions of $f$ 
and $g$. The biharmonic equation (\ref{eq:biharmonic}) is solved with boundary 
conditions $u(\pm 1) = u'(\pm 1) = 0$, where the Neumann conditions follow 
from the continuity equation. The boundary conditions for Eq. (\ref{eq:g}) are 
$g(\pm 1) = 0$. The variational formulation becomes: Find ${u} \in 
\N{V}_N$, ${g} \in \D{V}_N$ and ${f} \in \D{V}_N$ such that
\begin{align}
	\frac{\partial }{\partial t} \left< \nabla^2 u, \N{\psi}\right>_w &= 
	\left<h_u, \N{\psi} \right>_w + \nu \left<\nabla^4u, \N{\psi}\right>_w 
	&\forall \N{\psi} \in \N{V}_N, \label{eq:u1} \\
	\frac{\partial}{\partial t}\left<g, \D{\psi}\right>_w &= \left<h_g, 
	\D{\psi}\right>_w + \nu 
	\left<\nabla^2 g, \D{\psi}\right>_w &\forall \D{\psi} \in \D{V}_N ,
	\label{eq:g1} \\
	\left<f, \D{\psi}\right>_w &= \left<\frac{\partial u}{\partial x}, 
	\D{\psi}\right>_w &\forall \D{\psi} \in \D{V}_N. \label{eq:f1}
\end{align}
The two remaining velocity components are computed by projection to the 
Dirichlet space $\D{V}_N$: Find 
${v} \in \D{V}_N$ and $w \in \D{V}_N$ such that
\begin{align}
\left<f, \D{\psi}\right>_w &= \left<\frac{\partial v}{\partial y} + 
\frac{\partial w}{\partial z}, \D{\psi}\right>_w \, &\forall \D{\psi} \in 
\D{V}_N, \label{eq:f2} \\
\left<g, \D{\psi}\right>_w &= \left<\frac{\partial w}{\partial y}  - 
\frac{\partial v}{\partial z}, \D{\psi}\right>_w \, &\forall \D{\psi} \in 
\D{V}_N, \label{eq:g2}
\end{align}
which simplifies considerably because all the derivatives are in periodic 
directions. Written in spectral space Eqs. (\ref{eq:f2}) and (\ref{eq:g2}) 
become simply
\begin{align}
\hat{f}_{\bm{k}} &= \imath \underline{m}\, \hat{v}_{\bm{k}} + \imath 
\underline{n}\, \hat{w}_{\bm{k}} &\forall \bm{k} \in \D{K}_N^0, \label{eq:f3} \\
\hat{g}_{\bm{k}} &= \imath \underline{m}\, \hat{w}_{\bm{k}} - \imath 
\underline{n}\, \hat{v}_{\bm{k}} & \forall \bm{k} \in \D{K}_N^0, \label{eq:g3}
\end{align}
where $\D{K}_N^0$ is used to denote that these equations can be solved for all 
wavenumbers except $m=n=0$. For $m=n=0$ we solve instead the 
corresponding transformed momentum equations in $y$ and $z$ directions:
\begin{align}
\frac{\partial }{\partial t} \D{B}_{lq}\hat{v}(q, 0, 0, t) &= 
\D{B}_{lq}\hat{\mathcal{H}}_y(q, 0, 0, t) + \nu \D{A}_{lq} \hat{v}(q, 0, 0, t) 
& \forall\, l \in [0, 1, 
\ldots, N_x-2], \label{eq:v0}\\
\frac{\partial }{\partial t} \D{B}_{lq}\hat{w}(q, 0, 0, t) &= 
\D{B}_{lq}\hat{\mathcal{H}}_z(q, 0, 0, t) + \nu \D{A}_{lq} \hat{w}(q, 0, 0, t) 
& \forall\, l \in [0, 1, \ldots, N_x-2].\label{eq:w0}
\end{align}

Equations (\ref{eq:u1}-\ref{eq:f1}) are also solved in spectral space. 
Inserting for (\ref{eq:u_solx}) and a similar expansion for $g$ in $\D{V}_N$, 
the inner products required to solve Eqs. (\ref{eq:u1}-\ref{eq:f1}) are
\begin{align}
\left<\nabla^4u, \N{\psi}_{\bm{k}}\right>_w &= h\left[ \left( 
\N{\phi}_q^{''''}, 
\N{\phi}_l\right)_w -2(\underline{m}^2+\underline{n}^2) \left( \N{\phi}_q^{''}, 
\N{\phi}_l\right)_w + (\underline{m}^2+\underline{n}^2)^2\left( \N{\phi}_q, 
\N{\phi}_l\right)_w  \right] \hat{u}_q, \\
\left< \nabla^2 u, \N{\psi}_{\bm{k}}\right>_w &= h\left[\left( \N{\phi}_q^{''}, 
\N{\phi}_l\right)_w - (\underline{m}^2+\underline{n}^2)\left( \N{\phi}_q, 
\N{\phi}_l \right)_w \right] \hat{u}_q, \\
\left< \nabla^2 g, \D{\psi}_{\bm{k}}\right>_w &= h\left[\left( \D{\phi}_q^{''}, 
\D{\phi}_l\right)_w - (\underline{m}^2+\underline{n}^2)\left( \D{\phi}_q, 
\D{\phi}_l \right)_w \right] \hat{g}_q, \\
\left<\frac{\partial u}{\partial x}, \D{\psi}_{\bm{k}}\right>_w &=
h\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w \hat{u}_q, \\
\left<h_u, \N{\psi}_{\bm{k}} \right>_w &= 
h\left[-(\underline{m}^2+\underline{n}^2) 
\N{\mathcal{S}}(\mathcal{H}_x) - \imath \underline{m}\left(\D{\phi}_q^{'}, 
\N{\phi}_l \right)_w \hat{\mathcal{H}}_{y, q} - \imath 
\underline{n}\left(\D{\phi}_q^{'}, \N{\phi}_l \right)_w 
\hat{\mathcal{H}}_{z, q}\right], 
\\
\left< h_g, \D{\psi}_{\bm{k}} \right>_w &= h\left[ \imath \underline{m}\, \D{B} 
\hat{\mathcal{H}}_z - \imath \underline{n}\, \D{B} \hat{\mathcal{H}}_{y} 
\right],
\end{align}
where for brevity in notation (as before) it is understood that the matrices 
act along the first dimension of the transformed variables, i.e.,  $ 
\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w \hat{u}_q$ is short for the matrix 
vector product $ \sum_{q=0}^{N_x-4}\left( \N{\phi}_q^{'}, \D{\phi}_l\right)_w 
\hat{u}(q, {m}, {n}, t)$. The inner products are used to set up linear systems 
of equations for the inhomogeneous wall normal direction. All inner products 
$\left(\cdot, \cdot\right)_w$ can be represented by sparse matrices.
%\begin{align}
%\frac{\partial }{\partial t} \left( \N{\psi}_j^{''}, \N{\psi}_k\right)_w 
%\hat{u}_j &=  \left(h_u, \N{\psi}_k \right)_w + \nu \left(\N{\psi}^{''''}_j, 
%\N{\psi}_k\right)_w \hat{u}_j, \, &\forall \N{\psi}_k \in \N{V}_N, \\
%\frac{\partial}{\partial t}\left(\D{\psi}_j, \D{\psi}_k\right)_w \hat{g}_j &= 
%\left(h_g, \D{\psi}_k\right)_w + \nu \left(\D{\psi}_j^{''}, 
%\D{\psi}_k\right)_w 
%\hat{g}_j, \, &\forall \D{\psi}_k \in \D{V}_N \\
%\left(\D{\psi}_j, \D{\psi}_k\right)_w \hat{f}_j &= \left(\N{\psi}_j^{'}, 
%\D{\psi}_k\right)_w \hat{u}_j, \, &\forall \D{\psi}_k \in \D{V}_N.
%\end{align}
The mass matrices, stiffness matrices $\N{A}_{lq} = \left( \N{\phi}_q^{''}, 
\N{\phi}_l\right)_w, 
\D{A}_{lq} = \left( \D{\phi}_q^{''}, \D{\phi}_l\right)_w 
$ and $\N{Q}_{lq} = \left(\N{\phi}^{''''}_q, \N{\phi}_l\right)_w$ are given by 
Shen, whereas $\left(\N{\phi}_q^{'}, \D{\phi}_l\right)_w$ and 
$\left(\D{\phi}_q^{'}, \N{\phi}_l\right)_w$ are given in the 
appendix.
